NMRlipids IV: Headgroup \& glycerol backbone structures, and cation binding in bilayers with PS lipids

### 2018-08-16 Manuscript/manuscriptPS.tex
Update the text

    DOPS/Na$^+$  & GROMOS-CKP1 \cite{??} \todoi{Correct citation(s) for CKP.} & 128 & 4480 & 0  & 303  & 500 & 100 & \cite{ckp1DOPS303K} \\
    DOPS/Na$^+$  & GROMOS-CKP2 \cite{??} \todoi{Correct citation(s) for CKP.} & 128 & 4480 & 0  & 303  & 500 & 100 & \cite{ckp2DOPS303K} \\

@@

Before using the PC headgroup order parameters to quantify the ion binding
affinity, it is important to quantify the response of the headgroup order
parameters to the known amount of bound charge~\cite{catte16,melcr18}.
This can be done using the experimental data from the mixtures of
monovalent cationic surfactants (dihexadecyldimethylammonium) and POPC~\cite{scherer89,melcr18},
see section \ref{electrometerCALIBRATION} in the supplementary information.
In this work, we also quantify the response of PC headgroup order parameters
to the negatively charged PS headgroups, which also follows the electrometer
concept in the experiments~\cite{scherer87},
see section \ref{electrometerFORmixtures} in the supplementary information.

@@

The headgroup and glycerol backbone order parameters of 
POPS measured in this work are in good agreement with the previously reported
values from $^2$H NMR experiments of DOPS \cite{browning80} (Fig. \ref{HGorderParameters}).
The $\beta$-carbon order parameter is significantly more negative and $\alpha$-carbon
experiences a significant forking in PS headgroup when compared with
the values previously measured for POPC \cite{ferreira13} (Fig. \ref{HGorderParameters}).  
These features have been intepreted to arise from a rigid PS headgroup
conformation, stabilized by hydrogen bonds or electrostatic
interactions \cite{browning80,buldt81}, but detailed structrural interpretation is not
available. 
%The detailed structural differences between the headgroups is, however, not known.
\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/PCPScomp.pdf}
  \caption{\label{HGorderParameters}
    Headgroup and glycerol backbone order parameters of POPS measured in this work compared
    with the values from DOPS ($^2$H NMR, 0.1M of NaCl) \cite{browning80} and 
    POPC  ($^{13}$C NMR) \cite{ferreira13} experiments. Signs of the PS order parameters
    are measured in this work. Signs of the PC order parameters are measured in Ref.~\cite{ferreira16}.
  }
\end{figure}

@@

As previously observed for PC lipids, the order parameters and structures
of headgroup and glycerol backbone show wide variety between different simulation
models for PS lipids (Figs. \ref{HGorderParametersPS} and \ref{HGstructuresPS}).
The models for PS lipids perform generally less well than for
PC lipids in the previous study
(Figs. 2 and 4 in Ref.~\cite{botan15} vs. Figs.~\ref{HGorderParametersPS} and \ref{comparisonTablePS}).
Therefore, interpretation of structural differences between PC and PS headgroups
from simulations is not straightforward. The two best performing models,
Slipids and CHARMM36, reproduce the larger forking of the $\alpha$-carbon 
and the Slipids model reproduces also the lower of the $\beta$-carbon order parameter
when comparing the PS results to PC (Fig. 2 in Ref.~\citenum{botan15} vs. Fig. \ref{HGorderParametersPS}).
Interestingly, the dihedral angle distributions 
%of CHARMM36 and Slipids in 
in these two models share significant similarities in the headgroup region (Fig.~\ref{dihedralsHG}). 
However, the glycerol backbone order parameters in Slipids model
significantly differ from CHARMM36 results and experiments (Fig. \ref{HGorderParametersPS}),
which can be related to the differences in the dihedral angle distributions 
of C1-C2-C3-O31 and C2-C3-O31-C31 (Fig.~\ref{dihedralsGLY}). 
Similar difference was previously observed for PC lipids \cite{botan15}
and the conformational differences are illustrated in Figure \ref{glycerol_buslaev}.

%\todo{Discussion is to be finished.
%  One possible conclusion could be the following:
%  The main differences between the models in the headgroup region are observed
%  for dihedrals C12-C11-O12-P and C11-C12-C13-O1A.
%  CHARMM36, CHARMM36UA and Slipids give very similar results to the
%  dihedral C11-C12-C13-O1A, which is close to the $\beta$-carbon.
%  The order parameters of $\beta$-carbons for these three models
%  are in best argeement with the experiments in figure \ref{HGorderParametersPS}.
%  On the other hand, Gromos-CKP models give better order parameters
%  for $\alpha$-carbon than Slipids, CHARMM36 or CHARMM36UA. 
%  In conclusion, the suggestion would be that the single peak for
%  observed at 120 degrees in CHARMMs and Slipids would
%  be more realistic for C11-C12-C13-O1A dihedral, while the single peak at 180 degrees observed
%  in CKP models and in Berger would be most realistic for C12-C11-O12-P dihedral.
%}

@@

%
%The total deviation from the experiments ($\Sigma$ in Fig. \ref{comparisonTablePS})
%for the best performing CHARMM36 model is larger for PS lipids (8)
%than for PC (3) \cite{botan15}.
%Therefore the interpretation of structural details of PS headgroup or
%differences between PC and PS lipids is very challenging 
%from the current MD simulation models.

>>>>

    DOPS/Na$^+$  & GROMOS-CKPM \cite{??} \todoi{Correct citation(s) for CKP.} & 128 & 4480 & 0  & 303  & 500 & 100 & \cite{ckp1DOPS303K} \\
    DOPS/Na$^+$  & GROMOS-CKP  \cite{??} \todoi{Correct citation(s) for CKP.} & 128 & 4480 & 0  & 303  & 500 & 100 & \cite{ckp2DOPS303K} \\

@@

Before using the PC headgroup order parameters to quantify the ion binding
affinity, it is important to quantify their response to the known amount of bound charge~\cite{catte16,melcr18}.
This can be done using the experimental data from the mixtures of
monovalent cationic surfactants (dihexadecyldimethylammonium) and POPC~\cite{scherer89,melcr18},
see section \ref{electrometerCALIBRATION} in the supplementary information.
In this work, we also quantify the response of PC headgroup order parameters
to the negatively charged PS headgroups, which also follows the electrometer
concept in the experiments~\cite{scherer87},
see section \ref{electrometerFORmixtures} in the supplementary information.

@@

The headgroup and glycerol backbone order parameters of 
POPS measured in this work are in good agreement with the previously reported
values from $^2$H NMR experiments of DOPS \cite{browning80} (Fig. \ref{HGorderParameters}).
When compared with the previously measured values for POPC \cite{ferreira13} (Fig. \ref{HGorderParameters}),
the $\beta$-carbon order parameter is significantly more negative and $\alpha$-carbon
experiences a significant forking in PS headgroup.
These features have been intepreted to arise from a rigid PS headgroup
conformation, stabilized by hydrogen bonds or electrostatic
interactions \cite{browning80,buldt81}, but detailed structrural interpretation is not
available. 
%The detailed structural differences between the headgroups is, however, not known.
\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/PCPScomp.pdf}
  \caption{\label{HGorderParameters}
    Headgroup and glycerol backbone order parameters of POPS measured in this work compared
    with the values from DOPS ($^2$H NMR, 0.1M of NaCl) \cite{browning80} and 
    POPC  ($^{13}$C NMR) \cite{ferreira13} experiments. Signs of the PS order parameters
    are measured in this work. Signs of the PC order parameters are measured in Ref.~\cite{ferreira16}.
  }
\end{figure}

@@

The headgroup and glycerol backbone of PS lipids show wide variety between different simulation
models in the order parameters and structures (Figs. \ref{HGorderParametersPS} and \ref{HGstructuresPS}),
as previously observed also for PC lipids. The models for PS lipids perform generally less well than for
PC lipids in the previous study
(Figs. 2 and 4 in Ref.~\cite{botan15} vs. Figs.~\ref{HGorderParametersPS} and \ref{comparisonTablePS}).
Therefore, interpretation of structural differences between PC and PS headgroups
from simulations is not straightforward.

The best performing models,
Slipids, CHARMM36 and CHARMM36ua, reproduce the larger forking of the $\alpha$-carbon 
and the Slipids model reproduces also the lower of the $\beta$-carbon order parameter
when comparing the PS results to PC (Fig. 2 in Ref.~\citenum{botan15} vs. Fig. \ref{HGorderParametersPS}).
Interestingly, the headgroup structures of these models share significant similarities
and are distinct from other models (Figs.~\ref{HGstructuresPS} and \ref{dihedralsHG}). 
The distribution of the dihedral angle for C$_\alpha$-C$_\beta$-C$_\gamma$-O$_\gamma$ shows
a single narrow maximum close to 120$^{\circ}$ in the best three models, while other models give several maxima
in different locations (Fig. \ref{dihedralsHG}). The restricted motion is also
visible in the sampled conformations (Fig.~\ref{HGstructuresPS}) and might
manifest the increased rigidity anticipated from the early experimental studies \cite{browning80,buldt81}.
Another distinct feature in the three best performing models is the distribution
of N-C$_\beta$-C$_\alpha$-O$_\alpha$ dihedral, which gives a large maxima close to 300$^{\circ}$
and a smaller maxima close to 60$^{\circ}$. This is also visible in the sampled
structures (Fig.~\ref{HGstructuresPS}).
\todo{There seems to be discrepancy between structures and dihedral distributions for CHARMM36ua}.
The glycerol backbone order parameters of C$_2$ and C$_3$ from Slipids simulations
differ significantly from the other simulation results and experiments (Fig. \ref{HGorderParametersPS}),
as observed previously also for PC lipids \cite{botan15}.
The origin of this difference is more difficult to track without more elaborate analysis,
because different models show very complicated patterns of distinct structures
in the glycerol backbone region (Figs.~\ref{HGstructuresPS} and \ref{dihedralsHG}).

%which can be related to the differences in the dihedral angle distributions 
%of C1-C2-C3-O31 and C2-C3-O31-C31 (Fig.~\ref{dihedralsGLY}).

%  One possible conclusion could be the following:
%  The main differences between the models in the headgroup region are observed
%  for dihedrals C12-C11-O12-P and C11-C12-C13-O1A.
%  CHARMM36, CHARMM36UA and Slipids give very similar results to the
%  dihedral C11-C12-C13-O1A, which is close to the $\beta$-carbon.
%  The order parameters of $\beta$-carbons for these three models
%  are in best argeement with the experiments in figure \ref{HGorderParametersPS}.
%  On the other hand, Gromos-CKP models give better order parameters
%  for $\alpha$-carbon than Slipids, CHARMM36 or CHARMM36UA. 
%  In conclusion, the suggestion would be that the single peak for
%  observed at 120 degrees in CHARMMs and Slipids would
%  be more realistic for C11-C12-C13-O1A dihedral, while the single peak at 180 degrees observed
%  in CKP models and in Berger would be most realistic for C12-C11-O12-P dihedral.
%}

@@

%
%The total deviation from the experiments ($\Sigma$ in Fig. \ref{comparisonTablePS})
%for the best performing CHARMM36 model is larger for PS lipids (8)
%than for PC (3) \cite{botan15}.
